Optimised Cost Considering Huffman Code for Biological Data Compression

### 2015-01-30 Article/Article.tex
data set new

\section{Results And Discussion}
\label{sec4}
The approach has been evaluated with different real biological data (genomes), these genomes from The National Center for Biotechnology Information (NCBI) available on ( http://www.ncbi.nlm.nih.gov) []. In table () the different data set are described with the size of each of them and the references on the biological data bank.
\begin{table}
\small
\label{datas}
\caption{Dataset description}
\begin{center}
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (bp)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &6,988,302 & CP009496  \\\hline

@@

\multirow{2}{*}{Genome 14}&Pseudomonas aeruginosa& \multirow{2}{*}{6,591,121  } &\multirow{2}{*}{ CP006982  }\\    &LES400&\\
\hline
\bottomrule
\end{tabular}
\end{center}
\end{table} 
The table 3 present the results founded by the basic Huffman code, cost considering algorithm and optimised cost considering algorithm. The result show that the number of bits of Huffman algorithm is the minimum number among the other algorithm but the cost is  very high. The cost considering algorithm improve the quality of the generated codes in terms of cost but the number of bits. The optimised cost considering algorithm try to find the best allocation of codes to frequencies while the cost constraint respected. The table 4 present the best founded number of bits with different penalty, for each genome we find the max number of useful penalty, after this value, increasing the penalty are in-useful (See figure 3), the number of bits achieve the minimum number but the cost stop decreasing (point 1 in figure 3) and this number of bits stabilized while the cost still increasing until it stabilized also (point 2 figure 3), after this point the cost and number of bits are stabilized.
\begin{table}
\small
\label{datas}
\caption{GABM for comparison for cost and number of bits of different approaches with different datasets}
\begin{center}
\begin{tabular}{c | c c | c c | c c}
\toprule
$\textbf{Data sets}$ & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA}\\\hline
 & $\textbf{Cost}$	& $\textbf{Bits}$ &	$\textbf{Cost}$&	$\textbf{Bits	}$&$\textbf{Cost}$&	$\textbf{Bits}$
\\\hline
Genome 1&76787151  &	37256819	&67416213 & 41332479	& 67416213 & 40503061 \\\hline
Genome 2& 100425402 &  	48778740	& 88430665	& 54662727	& 88430665&  53341739 \\\hline
Genome 3&75940155  	&36860555	&66745619	&40927187 &	66745619 & 40182241\\\hline
Genome 4&103552729  	&50047265	&90821835&	56147481&	90821835  & 54319891\\\hline
Genome 5&82234926  &	39772838	&71963876	&44254676 &	71963876 & 43392820\\\hline
Genome 6&83454842  	&40370894	&73038795	&44864009	&73038795 & 44003571\\\hline
Genome 7&92539488  &	44977876&	81416359&	49790957	&81416359 & 49127335\\\hline
Genome 8&99613856 & 	48183688	&87102639&	53817307	&87102639&	52858893\\\hline
Genome 9&71876739 &	34874617&	62998800 &	38469200&	62998800 8&	37831964\\\hline
Genome 10&75324432 & 	36576034	&66084958&	40409906	&66084958&	39739522\\\hline
Genome 11&80766360  &	39092450&	70620666	&43119140&	70620666&	42359156 \\\hline
Genome 12&74560825  &	36179579&	65359604	&39926210	&65359604&	39221148\\\hline
Genome 13&71562941 & 	34737083	&62758225 &	38299629&	62758225 &	37809443\\\hline
Genome 14&78261299 & 	37843793&	68354090&	41816580&	68354090&	 41050420\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} \\\hline
Mycobacterium smegmatis &	 70760174 &37809636& 4\\\hline
Amycolatopsis benzoatilytica &	92010490& 49654132& 3\\\hline
Mycobacterium rhodesiae&	69421783& 37563017 &3\\\hline
Streptomyces bottropensis&	96294638& 50330594& 5\\\hline
Mycobacterium smegmatis. MC2&	74855668 &40380464& 5\\\hline
Mycobacterium smegmatis MKD8&	76738330& 40782100 &4\\\hline
Bradyrhizobium WSM471&	84667949 &45961149& 3\\\hline
Amycolatopsis thermoflava&	92416243 &48336727& 5\\\hline
Bacillus thuringiensis serovar thuringiensis&	66145783& 35330943& 4\\\hline
Bacillus thuringiensis Bt407&	68751359& 37306497 &3\\\hline
Pseudomonas aeruginosa 9BR&	72737300 &40264790& 2\\\hline
Bacillus thuringiensis serovar berliner&	67981988 &36770118 &3\\\hline
Bacillus thuringiensis serovar pakistani&	65896779& 35069999 &4\\\hline
Pseudomonas aeruginosa LES400&	71762305 &38210997 &4\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimised cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
\section{Conclusion}
\label{sec5}
In this paper we have proposed a new approach for efficient data compression using Huffman code and optimised strategy. The new approach is divided into phases, firstly a cost considering Huffman algorithm are proposed which reduce the cost of the generated codes, these codes are secondary passed by the optimisation algorithm to reduce the global number of bits using a cost penalty.\\
The proposed approach is tested with biological genomic sequence and a performance comparison is made with the standard Huffman code and the cost considering without optimisation. Simulation results showed that the proposed approach is more robust and efficient compared to other competing algorithms because its penalty based optimisation strategy to search the best allocation of codes to different frequencies.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num} 
%%  \bibliography{<your bibdatabase>}

>>>>

\section{Results And Discussion}
\label{sec4}
The approach has been evaluated with different real genomic biological data, these genomes were downloaded from a recent version of The National Center for Biotechnology Information (NCBI) available on $(http://www.ncbi.nlm.nih.gov)$. We focused on the sequences alone, ignoring any header and any other exogenous information. In table () the different data set are described with the size of each of them and the references on the biological data bank.
\begin{table}
\small
\label{datas}
\caption{Dataset description}
\begin{center}
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MegaOctet)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &6,988,302 & CP009496  \\\hline

@@

\multirow{2}{*}{Genome 14}&Pseudomonas aeruginosa& \multirow{2}{*}{6,591,121  } &\multirow{2}{*}{ CP006982  }\\    &LES400&\\
\hline
\bottomrule
\end{tabular}
\end{center}
\end{table} 
\begin{table}
\small
\label{datas}
\caption{GABM for comparison for cost and number of bits of different approaches with different datasets}
\begin{center}
\begin{tabular}{c | c c | c c | c c}
\toprule
$\textbf{Data sets}$ & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA}\\\hline
 & $\textbf{Cost}$	& $\textbf{Bits}$ &	$\textbf{Cost}$&	$\textbf{Bits	}$&$\textbf{Cost}$&	$\textbf{Bits}$
\\\hline
Genome 1&76787151  &	37256819	&67416213 & 41332479	& 67416213 & 40503061 \\\hline
Genome 2& 100425402 &  	48778740	& 88430665	& 54662727	& 88430665&  53341739 \\\hline
Genome 3&75940155  	&36860555	&66745619	&40927187 &	66745619 & 40182241\\\hline
Genome 4&103552729  	&50047265	&90821835&	56147481&	90821835  & 54319891\\\hline
Genome 5&82234926  &	39772838	&71963876	&44254676 &	71963876 & 43392820\\\hline
Genome 6&83454842  	&40370894	&73038795	&44864009	&73038795 & 44003571\\\hline
Genome 7&92539488  &	44977876&	81416359&	49790957	&81416359 & 49127335\\\hline
Genome 8&99613856 & 	48183688	&87102639&	53817307	&87102639&	52858893\\\hline
Genome 9&71876739 &	34874617&	62998800 &	38469200&	62998800 8&	37831964\\\hline
Genome 10&75324432 & 	36576034	&66084958&	40409906	&66084958&	39739522\\\hline
Genome 11&80766360  &	39092450&	70620666	&43119140&	70620666&	42359156 \\\hline
Genome 12&74560825  &	36179579&	65359604	&39926210	&65359604&	39221148\\\hline
Genome 13&71562941 & 	34737083	&62758225 &	38299629&	62758225 &	37809443\\\hline
Genome 14&78261299 & 	37843793&	68354090&	41816580&	68354090&	 41050420\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\small
\label{datas}
\caption{influence of penalty on bit minimization}
\begin{center}
\begin{tabular}{c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Bits}$ &		\textbf{$\lambda (\%)$} \\\hline
Mycobacterium smegmatis &	 70760174 &37809636& 4\\\hline
Amycolatopsis benzoatilytica &	92010490& 49654132& 3\\\hline
Mycobacterium rhodesiae&	69421783& 37563017 &3\\\hline
Streptomyces bottropensis&	96294638& 50330594& 5\\\hline
Mycobacterium smegmatis. MC2&	74855668 &40380464& 5\\\hline
Mycobacterium smegmatis MKD8&	76738330& 40782100 &4\\\hline
Bradyrhizobium WSM471&	84667949 &45961149& 3\\\hline
Amycolatopsis thermoflava&	92416243 &48336727& 5\\\hline
Bacillus thuringiensis serovar thuringiensis&	66145783& 35330943& 4\\\hline
Bacillus thuringiensis Bt407&	68751359& 37306497 &3\\\hline
Pseudomonas aeruginosa 9BR&	72737300 &40264790& 2\\\hline
Bacillus thuringiensis serovar berliner&	67981988 &36770118 &3\\\hline
Bacillus thuringiensis serovar pakistani&	65896779& 35069999 &4\\\hline
Pseudomonas aeruginosa LES400&	71762305 &38210997 &4\\\hline
\bottomrule
\end{tabular}
\end{center}
\end{table}
\begin{figure}[h]
\begin{center}
\includegraphics{Images/Drawing4.jpg}
\caption{Convergence of Optimised cost considering Huffman algorithm}
\end{center}
\label{Fig5}
\end{figure}
The table 3 present the results founded by the basic Huffman code, cost considering algorithm and optimised cost considering algorithm. The result show that the number of bits of Huffman algorithm is the minimum number among the other algorithm but the cost is  very high. The cost considering algorithm improve the quality of the generated codes in terms of cost but the number of bits. The optimised cost considering algorithm try to find the best allocation of codes to frequencies while the cost constraint respected. The table 4 present the best founded number of bits with different penalty, for each genome we find the max number of useful penalty, after this value, increasing the penalty are in-useful (See figure 3), the number of bits achieve the minimum number but the cost stop decreasing (point 1 in figure 3) and this number of bits stabilized while the cost still increasing until it stabilized also (point 2 figure 3), after this point the cost and number of bits are stabilized.

\section{Conclusion}
\label{sec5}
In this paper we have proposed a new approach for efficient data compression using Huffman code and optimised strategy. The new approach is divided into phases, firstly a cost considering Huffman algorithm are proposed which reduce the cost of the generated codes, these codes are secondary passed by the optimisation algorithm to reduce the global number of bits using a cost penalty.\\
The proposed approach is tested with biological genomic sequence and a performance comparison is made with the standard Huffman code and the cost considering without optimisation. Simulation results showed that the proposed approach is more robust and efficient compared to other competing algorithms because its penalty based optimisation strategy to search the best allocation of codes to different frequencies.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix
%% \section{}
%% \label{}
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num} 
%%  \bibliography{<your bibdatabase>}
